---
title: "CaseStudy"
format: pdf
---

```{r}
library(tidyverse)
library(gridExtra)
library(fields)
```

```{r}
df <- read.csv("NystedFarms.csv")
```

```{r}
# Figure 2
ggplot() + 
  geom_point(aes(x=XPos, y=YPos), 
             colour='lightgrey', data=df) +
  geom_point(aes(x=XPos, y=YPos, size=log(Count)), 
             colour="#377eb8", data=subset(df, Count>0)) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank())+
  coord_equal()
```
Figure 1.2: Transect locations (grey lines) and locations with non-zero counts (blue dots). The size of the blue dots are scaled based on the associated count at each location and are shown along the transects.



```{r}
# Figure 3
# Compute CIs for each construction phase using normal approx.
alpha <- 0.05
z_alpha_2 <- abs(qnorm(alpha/2)) # z_{alpha/2}
stats <- df |>
  group_by(Phase) |>
  summarise(Mean=mean(Count),
            Var=var(Count),
            N=n(),
            CI=z_alpha_2*sqrt(Var/N))

# Plot mean and 95% CIs
ggplot(stats, aes(x=Phase, y=Mean, colour=Phase)) +
  geom_errorbar(aes(ymin=Mean-CI, ymax=Mean+CI), size=1) +
  geom_point(size=4) +
  geom_text(aes(x=Phase, y=Mean+1.2*CI, label=paste0('N=', N))) +
  theme_bw() +
  ylab("Counts")
```
Figure 1.3: Mean counts per unit area (with 95% CIs) across phases.

```{r}
# Compute CIs for density using normal approx.
alpha <- 0.05
z_alpha_2 <- abs(qnorm(alpha/2)) # z_{alpha/2}
stats <- df %>%
  group_by(YearMonth, Phase) %>%
  summarise(Mean=mean(Count/Area),
            Var=var(Count/Area),
            N=n(),
            CI=z_alpha_2*sqrt(Var/N))

# Plot mean and 95% CIs
ggplot(stats, aes(x=YearMonth, y=Mean, colour=Phase)) +
  geom_errorbar(aes(ymin=Mean-CI, ymax=Mean+CI), size=1) +
  geom_point(size=2) +
  coord_flip() +
  theme(axis.text.x=element_text(angle=90)) +
  theme_bw() +
  ylab("Abundance per unit area") +
  xlab("Year/Month")
```
Figure 1.4: Average counts per unit area (with95% CIs) across years and months (with phases indicated).

```{r}
# Create list of plots of density against other covariates
p <- list()
for (covariate in c("XPos", "YPos", "Depth", "DistCoast"))
{
  p[[covariate]] <- ggplot(df, aes_string(x=covariate, 
                                          y="Count/Area")) + 
    geom_point(alpha=1/2) +
    ylab("Abundance per unit area")    
}

# Arrange ggplots in a grid
grid.arrange(grobs=p)
```

```{r}
quilt.plot(df$XPos, df$YPos, df$DistCoast, nx=35, ny=35,
           xlab="XPos", ylab="YPos", asp=1)
```
Figure 1.6: Distance from coast across the surveyed area. The colours shown in each grid cell represent the average distance from coast in that cell.


```{r}
quilt.plot(df$XPos, df$YPos, df$Depth, nx=35, ny=35,
           xlab="XPos", ylab="YPos", asp=1)
```
Figure 1.7: Depth across the surveyed area. The colours shown in each grid cell represent the average depth in that cell.


```{r}
par(mfrow=c(2,2))
quilt.plot(df$XPos[df$Phase=="A"], df$YPos[df$Phase=="A"], df$Count[df$Phase=='A']/df$Area[df$Phase=='A'], pch=20, 
           main="Phase A", xlab="X co-ordinate", ylab="Y co-ordinate", 
           nrow=20, ncol=20, asp=1)
quilt.plot(df$XPos[df$Phase=="B"], df$YPos[df$Phase=="B"], df$Count[df$Phase=='B']/df$Area[df$Phase=='B'], pch=20, 
           main="Phase B", xlab="X co-ordinate", ylab="Y co-ordinate", 
           nrow=20, ncol=20, asp=1)
quilt.plot(df$XPos[df$Phase=="C"], df$YPos[df$Phase=="C"], df$Count[df$Phase=='C']/df$Area[df$Phase=='C'], pch=20, 
           main="Phase C", xlab="X co-ordinate", ylab="Y co-ordinate", 
           nrow=20, ncol=20, asp=1)
par(mfrow=c(1,1))
```

```{r}
ggplot() + 
  geom_point(aes(x=XPos, y=YPos), colour='lightgrey', data=df) +
  geom_point(aes(x=XPos, y=YPos, size=log(Count/Area)), 
             colour="#377eb8", data=subset(df, Count>0)) +
  scale_size(range=c(1,3)) +
  theme_bw() +
  theme(panel.grid.major=element_blank(), 
        panel.grid.minor=element_blank()) +
  coord_equal() +
  facet_wrap(~Phase,nrow = 3)
```






